---
title: "How to Run R Codes"
author: "Seung Hoon Lee"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  xaringan::moon_reader:
    css: [default, c407_slides.css]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set()
```

## Two ways to run R codes

Console

R Script

---

## Console

Type the following codes in Console panel of RStudio, line by line.

```{r, eval=FALSE}
mtcars
```

```{r, eval=FALSE}
mean(mtcars$mpg)
```

```{r, eval=FALSE}
plot(mtcars$hp, mtcars$mpg)
```

Note that R shows the results in the same Console or in the Plots pane.

---

## R Script

Console: hard to reproduce or revise your analysis later. 

R Script allows you to save the whole codes as a file.

---

Click 'File - New File - R Script`.

In the new Script editor, type the following codes. 

- Press 'ctrl-enter' instead of enter key, *at the end of each line*:

```{r, eval=FALSE}
mtcars
```

```{r, eval=FALSE}
mean(mtcars$mpg)
```

```{r, eval=FALSE}
plot(mtcars$hp, mtcars$mpg)
```

Save the program as a file in 'Temp' folder, by clicking 'File - Save'.

---

You can run the whole program by pressing 'ctrl-shift-enter' key.

- Press 'ctrl-shft-F10' to restart R.

- Go to the Script editor, and press 'ctrl-shift-enter'.

---

Use both Console and R Script for an analysis.

- Include the codes you want to keep, in your R Script.

- Run the codes you do not want to keep, in Console.

    - Package installation.
    - Quick checks
    
