---
title: "Test Drive"
author: "Sanghoon Lee"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  xaringan::moon_reader:
    css: [default, c407_slides.css]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Go to the 'C407' folder.

Open 'C407.Rproj' file.

- Project file. (See [here](https://r4ds.had.co.nz/workflow-projects.html#rstudio-projects) for detail.)

- Sets the working folder to the folder where the file is located.

--

Create a new RScript.

- Click 'File - New File - R Script'.

- Save it in the 'Code' folder, as 'Test drive.R'.

---

In Console, type

```{r eval=FALSE}
install.packages("tidyverse")
```

--

In the RScript editor, 

```{r}
library(tidyverse)
library(lubridate)
bca <- read_rds("data/bca_sm.rds")
```

- At the end of each line, press *Ctrl + Enter*.
- The shortcut key for <- is *alt + -*.

---

In Console,
```{r eval=FALSE}
bca %>% View
```

- The shortcut key for %>% is *Ctrl + Shft + m*.

---

.scroll-output[
```{r echo=FALSE}
op <- options("width"=1000)
bca %>% head(3) %>% kableExtra::kable()
options(op)
```
]

---

Suppose you want to calculate average transaction price for each year. How would you do it?

---

In the RScript editor, 
```{r}
avg_price <- bca %>% 
	group_by(sale_year = year(sale_date)) %>% 
	summarise(avg_price = mean(price))
```

Press 'enter' at the end of the first two lines, and 'ctrl + enter' at the last line.

---

In the RScript editor, 
```{r eval=FALSE}
avg_price
```

---

.scroll-output[
```{r echo=FALSE}
avg_price %>% kableExtra::kable()
```
]

---

In the RScript editor, 
```{r eval=FALSE}
ggplot(avg_price, aes(sale_year, avg_price)) + 
	geom_line() + 
	geom_point() + 
	scale_y_continuous(labels = scales::comma) 
```

---

```{r echo=FALSE}
ggplot(avg_price, aes(sale_year, avg_price)) + geom_line() + geom_point() + scale_y_continuous(labels = scales::comma)
```

---

In the R Script editor,
```{r}
ggsave("Temp/avg_price.png")
```

- Check 'avg_price.png' is generated in 'C407/Temp' folder.