---
title: "Working Directory and Relative Path"
author: "Sanghoon Lee"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  xaringan::moon_reader:
    css: [default, c407_slides.css]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(kableExtra)
```

## Working Directory

The (current) working directory is the directory R is currently working in.

--

The working directory is the default location R uses when reading or writing a file.

--

When you open the 'c407' project file, it sets the working directory to be the one where the project file is located, i.e., 'c407' folder.

---

Open the 'c407' project file.

--

The top of Console pane shows the working directory.

--

Or you can type in Console
```{r eval=FALSE}
getwd()
```

--

Confirm that the directory is the one where the project file is located.

To see contents in the current working folder in 'Files' pane, click 'More - Go to Working Directory' in the pane.

---

The working directory is the default location R uses when reading or writing a file.

```{r}
library(tidyverse)
```

```{r eval=FALSE}
bca <- read_rds("bca_sm.rds")
```

--

The command above does not work, since R cannot find the file in the current working folder.

---

Instead, use
```{r}
bca <- read_rds("Data/bca_sm.rds")
```

--

Confirm in 'Files' pane that the 'bca_sm.rds' is in 'Data' subfolder from the working directory.

--

Note we use '/' rather than '\'.

--

This is called a relative path.

- A relative path is shorter than an absolute path.

- It is more portable.

---

```{r}
ggplot(mtcars) + geom_point(aes(mpg,hp))
```

---

The following saves the plot in 'c407', the current working folder.
```{r eval=FALSE}
ggsave("mpg_hp.png")
```

--

If you want to save the plot in 'Temp' subfolder, use
--
```{r}
ggsave("Temp/mpg_hp.png")
```

Confirm in 'Files' pane, that the files are created in the two folders.

---

## Task

Open '2016_census_pumf_ind.rds' in 'Data' subfolder, and names it to 'pumf'.

--

```{r}
pumf <- read_rds('Data/2016_census_pumf_ind.rds')
```

--

Browse the data using 'View()' function.

--

```{r eval=FALSE}
View(pumf)
```

---

.scroll-output[
```{r echo=FALSE}
pumf %>% head(3) %>% kable()
```
]

