---
title: "Data Structure and Basic Operations"
author: "Sanghoon Lee"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  xaringan::moon_reader:
    css: [default, c407_slides.css]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, comment=NA)
```

Open `C407` project.

```{r}
library(tidyverse)
```

```{r}
bca <- read_rds("data/bca_sm.rds")
```

---

## Vector and Data Frame

Data come as a data frame (or a tibble in Tidyverse).

--

Data is a collection of columns equal length vectors (i.e., columns).

```{r eval=FALSE}
View(bca)
```

---

.scroll-output[
```{r echo = FALSE}
bca %>% head(20) %>% kableExtra::kable()
```
]

---

## Vector

A vector is a sequence of values of the *same type* (e.g., numeric, character, date, ...).

--

You use 'c()' function to create a vector.

Numeric
```{r}
a <- c(0,1,2)
```

Character
```{r}
b <- c("A", "B", "C")
```

---

```{r}
a
```

```{r}
class(a)
```

```{r}
b
```

```{r}
class(b)
```

---

### Why the values of the same type?

Some operations work on only specific types.

```{r}
a * a
```

```{r eval=FALSE}
b * b
```

---

## Data Frame

You can create a data frame using 'data.frame()' function.
```{r}
df <-  data.frame(a,b)
```

```{r}
df
```

---

```{r}
data.frame(a=c(0,1,2), b=c("A", "B", "C"))
```

---

## Tibble

A tibble is an enhanced version of data frame used in Tidyverse.
```{r}
tb <- tibble(a, b)
```

```{r}
tb
```

---

## Task

Create the following tibble, and name it 'tb2'. Print it on screen.
```{r echo=FALSE}
tb2 <- tibble( name = c("a", "b", "c"), x=c(1,2,3), y=c(3,2,1)  )
tb2
```

--

```{r eval=FALSE}
tb2 <- tibble( name = c("a", "b", "c"), x=c(1,2,3), y=c(3,2,1)  )
tb2
```

---

## Common operations with a tibble or a data frame

You can pull out a column as a vector, using '$'.
```{r}
tb$a
```

--

```{r}
mean(tb$a)
```

---

Show column names as a vector
```{r}
names(tb)
```

---

Show a summary of each column
```{r}
summary(tb)
```

---

Show the number of rows

```{r}
nrow(tb)
```

---

Show the first two rows
```{r}
head(tb, 2)
```

---

Without the second argument, it shows 6 rows.
```{r}
head(bca)
```

---

.scroll-output[
```{r}
glimpse(bca)
```
]

---

## Task

How many rows does 'bca' data set have?

--

```{r}
nrow(bca)
```

---

Show the column names of the 'bca' dataset.

--

```{r}
names(bca)
```

---

What is the median of price variable(i.e., column) in the bca dataset?

--

```{r}
median(bca$price)
```

--

Alternatively, you can use 'summary()' function.
```{r eval=FALSE}
summary(bca)
```

---

What is the type of 'postalcode' column?

--

```{r}
class(bca$postalcode)
```

```{r eval=FALSE}
bca$postalcode
```

```{r echo=FALSE}
bca$postalcode %>% head(300)
```

---

What is the type of 'price' column?

--

```{r}
class(bca$price)
```

```{r eval=FALSE}
bca$price
```

```{r echo=FALSE}
bca$price %>% head(300)
```

---

What is the type of 'sale_date' column?

```{r}
class(bca$sale_date)
```

```{r eval=FALSE}
bca$sale_date
```

```{r echo=FALSE}
bca$sale_date %>% head(300)
```

