---
title: "dplyr 1"
subtitle: "i-Clicker quiz"
author: "Sanghoon Lee"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  xaringan::moon_reader:
    css: [default, c407_slides.css]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, comment=NA)
library(kableExtra)
```

```{r echo=FALSE}
library(tidyverse)
```

You want to transform the tibble in the left to the one in the right.
.pull-left[
```{r echo=FALSE}
tibble(a=c(2,1,3), b=c("a", "b", "c")) %>% kable(align = "c") %>% column_spec(1:2, width = "15em")
```
]
.pull-right[
```{r echo=FALSE}
tibble(a=c(2,1,3), b=c("a", "b", "c")) %>% mutate(c=2*a) %>% kable(align = "c") %>% column_spec(1:3, width = "15em")
```
]
Which function would you use?

<ol type="A">
  <li>select()</li>
  <li>choose()</li>
  <li>filter()</li>
  <li>create()</li>
  <li>mutate()</li>
</ol>

---

You want to transform the tibble in the left to the one in the right.
.pull-left[
```{r echo=FALSE}
tibble(a=c(2,1,3), b=c("a", "b", "c")) %>% kable(align = "c") %>% column_spec(1:2, width = "15em")
```
]
.pull-right[
```{r echo=FALSE}
tibble(a=c(2,1,3), b=c("a", "b", "c")) %>% filter(a!=2) %>% kable(align = "c") %>% column_spec(1:2, width = "15em")
```
]

Which function would you use?

<ol type="A">
  <li>select()</li>
  <li>choose()</li>
  <li>keep()</li>
  <li>filter()</li>
  <li>mutate()</li>
</ol>

---

You want to transform the tibble in the left to the one in the right.
.pull-left[
```{r echo=FALSE}
tibble(a=c(2,1,3), b=c("a", "b", "c")) %>% kable(align = "c") %>% column_spec(1:2, width = "15em")
```
]
.pull-right[
```{r echo=FALSE}
tibble(a=c(2,1,3), b=c("a", "b", "c")) %>% select(b) %>% kable(align = "c") %>% column_spec(1, width = "15em")
```
]

Which function would you use?

<ol type="A">
  <li>select()</li>
  <li>choose()</li>
  <li>keep()</li>
  <li>filter()</li>
  <li>mutate()</li>
</ol>

