---
title: "dplyr I"
author: "Sanghoon Lee"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  xaringan::moon_reader:
    css: [default, c407_slides.css]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, comment=NA)
library(kableExtra)
```

select()

filter()

arrange()

mutate()

---

```{r}
library(tidyverse)
```

```{r results='hide'}
tb <- tibble( a=c(2,1,3), b=c("a", "b", "c"), c=c("y","y","x") )
tb
```

```{r echo=FALSE}
tb %>% kable(align = "c") %>% column_spec(1:3, width = "5em")
```

---

## select()

```{r eval=FALSE}
? select
```

--

Choose variables from a tibble.

**Usage**

select(.data, ...)

... variables to select

---

.pull-left[
```{r}
tb
```
]

.pull-right[
```{r}
select(tb, a, b)
```
]

---

## Pipe %>% 

* `x %>% f(...)` is equivalent to `f(x,...)`.

--

.pull-left[
```{r}
select(tb, a, b)
```
]

.pull-right[
```{r}
tb %>% select(a,b)
```
]

---

## filter()

Return rows with matching conditions.

.pull-left[
```{r}
tb
```
]

.pull-right[
```{r}
filter(tb, a >= 2, c == "x")
```
]

---

* `x %>% f(...)` is equivalent to `f(x,...)`.

.pull-left[
```{r}
filter(tb, a >= 2, c == "x")
```
]

.pull-right[
```{r}
tb %>% filter( a >= 2, c == "x")
```
]

---

## Why Pipe %>% ?

--

Easier to read.
.pull-left[
```{r}
filter(tb, a >= 2, c == "x")
```
]

.pull-right[
```{r}
tb %>% filter( a >= 2, c == "x")
```
]

---

Chain multiple operations.

Suppose you want to do the following to tb sequently:

1. select columns a and c.
2. choose rows such that a>= 2 and c == "x".

--

.pull-left[
```{r}
temp <- select(tb, a, c)
filter(temp, a >= 2, c == "x")
```
]

--

.pull-right[
```{r}
tb %>% 
  select(a, c) %>% 
  filter(a >= 2, c == "x")
```
]

---

## arrange()

Sort rows by variables

.pull-left[
```{r}
tb
```
]

.pull-right[
```{r}
tb %>% arrange(c, a)
```
]

---

.pull-left[
```{r}
tb
```
]

.pull-right[
```{r}
tb %>% arrange(desc(c), a)
```
]

---

## mutate()

Create variables
--
.pull-left[
```{r}
tb
```
]
.pull-right[
```{r}
tb %>% mutate(d = 3*a, 
              e = paste0(b, c) )
```
]

---

## Task

From the bca_sm dataset, generate a tibble of housing transactions satisfying the following:

- Create a new column called 'sale_year' containing the year of sale. (Use 'year()' function in 'lubridate' package.)
- Sold in 2017.
- Condo or townhouse (i.e., "Strata Res" in act_use_cat)
- Price is greater than 5 million dollars.
- Contain only the following columns: roll_num, sale_year, price, neigh_label.
- Sorted in the descending order of price.

---

Restart R

--

Load libraries.

--

```{r}
library(lubridate)
library(tidyverse)
```

--

Load the bca_sm.rds.

--

```{r}
bca <- read_rds("Data/bca_sm.rds")
```

--

Browse data to identify relevant columns.

```{r eval=FALSE}
View(bca)
```

---

Create a new column called 'sale_year' containing the year of sale. (Use 'year()' function in 'lubridate' package.)

--

```{r}
temp <- bca %>% mutate(sale_year = year(sale_date))
```

--

Confirm that 'sale_year' column is created.
```{r eval=FALSE}
temp %>% View
```

---

- Sold in 2017.
- Condo or townhouse (i.e., "Strata Res" in act_use_cat)
- Price is greater than 5 million dollars.

--

```{r}
temp <- temp %>% filter(
  sale_year == 2017, 
  act_use_cat == "Strata Res", 
  price>5000000 )
```

--

Confirm that they are filtered correctly.
```{r eval=FALSE}
temp %>% View
```

---

Contain only the following columns: roll_num, sale_year, price, neigh_label.

```{r}
temp <- temp %>% select(roll_num, sale_year, price, neigh_label)
```

--

Confirm the result:
```{r}
temp
```

---

Sorted in the descending order of price.

--

```{r}
temp %>% arrange(desc(price))
```

---

```{r eval=FALSE}
temp <- bca %>% mutate(sale_year = year(sale_date))

temp <- temp %>% filter(
  sale_year == 2017, 
  act_use_cat == "Strata Res", 
  price>5000000 )

temp <- temp %>% 
  select(roll_num, sale_year, price, neigh_label)

temp %>% arrange(desc(price))
```

Try rewriting the codes using the pipe, without using a temp object!

---

```{r eval=FALSE}
bca %>% mutate(sale_year = year(sale_date)) %>% 
  filter(
    sale_year == 2017, 
    act_use_cat == "Strata Res", 
    price>5000000 ) %>% 
  select(roll_num, sale_year, price, neigh_label) %>% 
  arrange(desc(price))  
```

