---
title: "dplyr 2"
author: "Sanghoon Lee"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  xaringan::moon_reader:
    css: [default, c407_slides.css]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, comment=NA)
library(kableExtra)
```


The following code sums up the population sizes of counties by state. Which 

```{r eval=FALSE}
counties %>% 
  count(state, ___ = population, sort = TRUE )
```

Which functions do you use to 

<ol type="A">
  <li>sum</li>
  <li>wt</li>
  <li>weight</li>
  <li>size</li>
  <li>tot</li>
</ol>

---

The counties data set has unemployment rate and state name for each county. You want to calculate average unemployment rate of counties, by each state. Which functions would you use?

<ol type="A">
  <li>average_by()</li>
  <li>group_by(), calculate(), mean()</li>
  <li>group_by(), mutate(), mean()</li>
  <li>group_by(), summarise(), mean()</li>
  <li>mutate_by(), mean()</li>
</ol>

---

