---
title: "dplyr 2"
author: "Sanghoon Lee"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  xaringan::moon_reader:
    css: [default, c407_slides.css]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, comment=NA)
library(kableExtra)
```

count()

summarise()

group_by()

ungroup()

top_n()

---

```{r}
library(tidyverse)
```
.pull-left-2[
```{r}
counties <- tibble(
  county=c("a", "b", "c", "d", "e"), 
  pop=c(1,2,3,4,5), 
  income=c(14, 13, 12, 8, 10), 
  state=c("x","x","x", "y","y"))
```
]
.pull-right-1[
```{r}
counties
```
]

---

### ? count()

Count observations by group

*Usage*
count(x, ..., wt = NULL,sort= FALSE)

- x	a tibble to count.
- ...	Variables to group by.
- wt (Optional) If omitted (and no variable named n exists in the data), will count the number of rows. If specified, will perform a "weighted" tally by summing the (non-missing) values of variable wt. 
- sort	if TRUE will sort output in descending order of n

---

.pull-left[
```{r}
counties
```
]
.pull-right[
```{r}
count(counties)
```
```{r}
counties %>% count()
```
]
---

.pull-left-1[
```{r}
counties
```
]

.pull-right-2[
```{r}
counties %>% count(state)
```
```{r}
counties %>% count(state, wt=pop)
```
]

---

```{r}
counties %>% count(state, wt=pop)
```

```{r}
counties %>% count(state, wt=pop, sort=TRUE)
```

---

## Application

Using 2016 Census public use microfile, estimate population size by census metropolitan area (CMA). 

- Use "data/2016_census_pumf_ind.rds".
- The dataset has WEIGHT column for each individual included in the survey.
- Sorted in descending order of the population size.
- Remove observaions where CMA is "Other census metropolitan areas, census agglomerations and other geographies".
- Show only the six largest CMAs.

---

```{r}
pumf <- read_rds("data/2016_census_pumf_ind.rds")
```

```{r eval=FALSE}
pumf %>% select(CMA, WEIGHT) %>% View
```
.scroll-output[
```{r echo=FALSE}
pumf %>% select(CMA, WEIGHT) %>% head(20) %>% kable()
```
]

---

```{r}
pumf %>% 
  count(CMA, wt=WEIGHT, sort=TRUE) %>% 
  filter(CMA != "Other census metropolitan areas, census agglomerations and other geographies") %>% 
  head()
```

---

## summarise()

.pull-left-1[
```{r}
counties
```
]
.pull-right-2[
```{r}
counties %>% 
  summarise(avg_income = mean(income))
```
]

---
.pull-left-1[
```{r}
counties
```
]
.pull-right-2[
```{r}
counties %>% 
  summarise(avg_income = weighted.mean(income, w= pop))
```
]

---

```{r}
counties %>% summarise(
  total_pop=sum(pop), 
  avg_income = weighted.mean(income, w= pop)
  )
```

---

.pull-left[
```{r}
counties %>% summarise(n=n())
```

```{r}
counties %>% 
  summarise(total_pop=sum(pop))
```
]

--

.pull-right[
```{r}
counties %>% count()
```

```{r}
counties %>% count(wt=pop)
```
]

---

### Summary functions

- sum()
- mean()
- weighted.mean()
- median()
- min()
- max()
- n()

--

They all generate one element result.

---

### group_by() - summarise()

With group_by(), summarise() works on each subset defined by the group variables.

---

.pull-left-1[
```{r}
counties
```
]
.pull-right-2[
```{r}
counties %>% 
  group_by(state) %>% 
  summarise(pop = sum(pop))
```
]

---

```{r}
counties %>% group_by(state) %>% 
  summarise(avg_income = weighted.mean(income, w=pop))
```

---

.pull-left[
```{r}
counties
```
]

.pull-right[
```{r}
counties %>% group_by(state)
```
]

group_by() adds group variables. 

When summarise() sees group variables, it runs on each subset defined by the group variables.

---

### Ungrouping

Each time summarise() is used, the last group variables gets removed.
```{r}
counties %>% group_by(state, county) %>% 
  summarise(county_pop = sum(pop))
```

---

Use ungroup() to remove all groupings.

```{r}
counties %>% group_by(state, county) %>% ungroup()
```

---

### Application

Using 2016 Census public use microfile, estimate population size by census metropolitan area (CMA). Use group_by()-summarise().

- Use "data/2016_census_pumf_ind.rds".
- The dataset has WEIGHT column for each individual included in the survey.
- Sorted in descending order of the population size.
- Remove observaions where CMA is "Other census metropolitan areas, census agglomerations and other geographies".
- Show only the six largest CMAs.

---

```{r}
pumf %>% group_by(CMA) %>% 
  summarise(n = sum(WEIGHT)) %>% 
  filter(CMA != "Other census metropolitan areas, census agglomerations and other geographies") %>% 
  arrange(desc(n)) %>% 
  head()
```

---

### top_n()

Select top n rows by value

top_n(x, n, wt)

- x	a tbl() to filter
- n	number of rows to return for top_n().
- If x is grouped, this is the number (or fraction) of rows per group. Will include more rows if there are ties.
- If n is positive, selects the top rows. If negative, selects the bottom rows.
- wt	The variable to use for ordering.

---

.pull-left-1[
```{r}
counties
```
]
.pull-right-2[
```{r}
counties %>% top_n(3, pop)
```
]

---

.pull-left-1[
```{r}
counties
```
]
.pull-right-2[
```{r}
counties %>% group_by(state) %>% 
  top_n(1, income)
```
]

---

### Application

Using 2016 Census public use microfile (PUMF), estimate average individual annual total income (TOTINC) by census metropolitan area (CMA).  

- Use TOTINC column for individual annual total income. (See the User Guide in data folder for details.)
- Remove those observations whose TOTINC is either 88888888 (Not available) or 99999999 (Not applicable). (See the User Guide in data folder for details.)
- Use weighted.mean function with w=WEIGHT argument.
- Show only the ten highest income CMAs.

---

```{r eval=FALSE}
pumf %>% select(CMA, TOTINC, WEIGHT) %>% View()
```

.scroll-output[
```{r echo=FALSE}
pumf %>% select(CMA, TOTINC, WEIGHT) %>% head(30) %>% kable
```
]

---

```{r}
pumf %>% 
  filter(TOTINC != 88888888, TOTINC != 99999999 ) %>% 
  group_by(CMA) %>% 
  summarise(avg_income = weighted.mean(TOTINC, w=WEIGHT)) %>% 
  top_n(10, avg_income)
```

