---
title: "dplyr 3"
author: "Sanghoon Lee"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  xaringan::moon_reader:
    css: [default, c407_slides.css]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, comment=NA)
library(kableExtra)
```

Select helpers

- :
- contains()
- starts_with()
- ends_with()
- last_col()
- -

rename()

transmute()

---

```{r}
library(tidyverse)
```

```{r}
tb <- tibble(
  a1 = c(1,1), a2 = c(1,2), 
  b1 = c(2,1), b2 = c(2,2), 
  ab1=c(3,1), ab2=c(3,2)
  )

tb
```

---

### Select helpers

.pull-left[
Using the select helper functions, select the following columns from tb.
- :
- contains()
- starts_with()
- ends_with()
- last_col()
]

.pull-right[
1. "a2", "b1", "b2", "ab1"
1. "a1", "a2", "b1", "b2", "ab1"
1. "a1",  "b1", "ab1"
1. "a1", "a2", "ab1", "ab2"
1. "b1", "ab1"
]

---

.pull-left[
```{r}
tb
```

- "a2", "b1", "b2", "ab1"
```{r eval=FALSE}
tb %>% select(a2:ab1)
```

- "a1", "a2", "b1", "b2", "ab1"
```{r eval=FALSE}
tb %>% select(-ab2)
```
]

.pull-right[

- "a1",  "b1", "ab1"
```{r eval=FALSE}
tb %>% select(ends_with("1"))
```

- "a1", "a2", "ab1", "ab2"
```{r eval=FALSE}
tb %>% select(starts_with("a"))
```

- "b1", "ab1"
```{r eval=FALSE}
tb %>% select(contains("b1"))
```
]

---

### Application

```{r}
pumf <- read_rds("data/2016_census_pumf_ind.rds")
```

Select the following variables: "CMA", "CMAUID", "PPSORT", "WEIGHT", "POB", "POBF", "POBM", and save the result as pob. Print the first 6 rows.

---

.scroll-output[
```{r}
pumf %>% names()
```
]
 
---

```{r}
pob <- pumf %>% select(CMA:WEIGHT, starts_with("POB"))
pob %>% head()
```

---

Estimate the number of people by POBF and POBM, sort the counts in the descending order, and save the result as pob_n. Print the top 20 rows.

---

```{r}
pob_n <- pob %>% 
  group_by(POBF, POBM) %>% 
  summarise(n=sum(WEIGHT)) %>% 
  arrange(desc(n))
```

--

```{r eval=FALSE}
pob_n <- pob %>% 
  count(POBF, POBM, wt=WEIGHT, sort=TRUE)
```

---

.scroll-output[
```{r}
pob_n %>% top_n(20)
```
]

---

Remove the observations where POBF and POBM are equal. Show the top 20 rows.

--

.scroll-output[
```{r}
pob_n %>% filter(POBF != POBM) %>% top_n(20)
```
]

---

### rename()

Rename a1 and a2 to c1 and c2 respectively

.pull-left[
```{r}
tb
```
]

.pull-right[
```{r}
tb %>% rename(c1=a1, c2=a2)
```
]

---

You can also rename while selecting or grouping.

.pull-left[
```{r}
tb %>% select(c1=a1, c2=a2)
```
]

.pull-right[
```{r}
tb %>% group_by(c1=a1) %>% 
  summarise(avg_a2 = mean(a2))
```
]

But do not use select() or group_by() just for renaming columns.

---

### transmute()

mutate() + select()

.pull-left[
```{r}
tb
```
]

.pull-right[
```{r}
tb %>% transmute(a1, a3=a1+a2)
```
]

---

### Application

```{r}
bca <- read_rds("data/bca_sm.rds")
```

Create three columns, sale_wday, sale_month, sale_year, that contain the weekday, month and year of sale. (Use wday(), month(), year() in lubridate package. Use label=TRUE as the second argument in wday() and month().)

Keep only the four columns: price and the new sale_wday, sale_month, and sale_year columns, and save it as temp.

Remove observations with the sale_year == 2018 in temp. (The bca data end in Aug 8 2018.)

Using temp, calculate the following:

- the number of sales and average price by each weekday.
- the number of sales and average price by month.
- the number of sales and average price by year.

---

```{r}
library(lubridate)

temp <- bca %>% 
  transmute(
    price, 
    sale_wday = wday(sale_date, label=TRUE),
    sale_month = month(sale_date, label=TRUE),
    sale_year = year(sale_date)
            ) %>% 
  filter(sale_year != 2018)
```

---

```{r}
temp %>% group_by(sale_wday) %>% 
  summarise(n = n(), avg_price = mean(price))
```

---

```{r}
temp %>% group_by(sale_month) %>% 
  summarise(n = n(), avg_price = mean(price))
```

---

```{r}
temp %>% group_by(sale_year) %>% 
  summarise(n = n(), avg_price = mean(price))
```
