---
title: "ggplot2 - 1"
author: "Sanghoon Lee"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  xaringan::moon_reader:
    css: [default, c407_slides.css]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, comment=NA)
options(tibble.print_max = Inf)
library(kableExtra)
```

ggplot2 grammar

geom_point()

geom_smooth()

labs()

theme()

Saving a part of the whole plot

---

```{r eval=FALSE}
install.packages("gapminder")
```
Need to install gapminder package, just once.

```{r}
library(gapminder)
library(tidyverse)
```

No need to load ggplot2 package in addition to tidyverse.

```{r}
gapminder_2007 <- gapminder %>% filter(year == 2007)
```

```{r eval=FALSE}
gapminder_2007
```

---

.scroll-output[
```{r echo=FALSE}
gapminder_2007 %>% kable()
```
]

---

### Basic ggplot structure

.small[
```{r}
ggplot() + 
  geom_point(data=gapminder_2007, mapping=aes(x=gdpPercap, y=lifeExp))
```
]

---

### Different geom

.small[
```{r}
ggplot() + 
  geom_smooth(data=gapminder_2007, mapping=aes(x=gdpPercap, y=lifeExp))
```
]

---

### Overlaying multiple geoms

.small[
```{r}
ggplot() + 
  geom_smooth(data=gapminder_2007, mapping=aes(x=gdpPercap, y=lifeExp)) +
  geom_point(data=gapminder_2007, mapping=aes(x=gdpPercap, y=lifeExp))
```
]

---

### Global option

.small[
```{r eval=FALSE}
ggplot() + 
  geom_smooth(data=gapminder_2007, mapping=aes(x=gdpPercap, y=lifeExp)) +
  geom_point(data=gapminder_2007, mapping=aes(x=gdpPercap, y=lifeExp)) 
```
]

--

```{r eval=TRUE, fig.height=4.5}
ggplot(data=gapminder_2007, mapping=aes(x=gdpPercap, y=lifeExp)) + 
  geom_smooth() +  geom_point()
```

---

### Matching arguments by position

[? `ggplot`](https://ggplot2.tidyverse.org/reference/ggplot.html)

[? `aes`](https://ggplot2.tidyverse.org/reference/aes.html)

```{r eval=FALSE}
ggplot(data=gapminder_2007, mapping=aes(x=gdpPercap, y=lifeExp)) + 
  geom_smooth() +  geom_point()
```

```{r}
ggplot(gapminder_2007, aes(gdpPercap, lifeExp ) ) + 
  geom_smooth() + geom_point()
```

---

[`? geom_point`](https://ggplot2.tidyverse.org/reference/geom_point.html)

```{r}
ggplot(gapminder_2007) + 
  geom_point(aes(gdpPercap, lifeExp))
```

---

### Aesthetic mappings

```{r}
ggplot(gapminder_2007) + 
  geom_point(aes(gdpPercap, lifeExp, color=continent))
```

---

```{r}
ggplot(gapminder_2007) + 
  geom_point(aes(gdpPercap, lifeExp, shape=continent))
```

---

.small[
```{r}
ggplot(gapminder_2007) + 
  geom_point(aes(gdpPercap, lifeExp, shape=continent, color=continent))
```
]

---

### labs()

[? `labs`](https://ggplot2.tidyverse.org/reference/labs.html)

labs(..., title = waiver(), subtitle = waiver(), caption = waiver(),
  tag = waiver()) 
  - ...	A list of new name-value pairs. The name should be an aesthetic.

.small[
```{r eval=FALSE}
ggplot(gapminder_2007) + 
  geom_point(aes(gdpPercap, lifeExp, color=continent)) +
  labs(x= "GDP per capita", y="Life Expectancy", color="Continent", 
       title = "Income vs Longevity", caption = "Source: Gapminder dataset")
```
]

---

.small[
```{r}
ggplot(gapminder_2007) + 
  geom_point(aes(gdpPercap, lifeExp, color=continent)) +
  labs(x= "GDP per capita", y="Life Expectancy", color="Continent", 
       title = "Income vs Longevity", caption = "Source: Gapminder dataset")
```
]

---

### theme()

.small[
```{r}
ggplot(gapminder_2007) + 
  geom_point(aes(gdpPercap, lifeExp, color=continent)) +
  labs(x= "GDP per capita", y="Life Expectancy", color="Continent", 
       title = "Income vs Longevity", caption = "Source: Gapminder dataset")+
  theme_bw()
```
]

---

### Saving a part of ggplot code

Try

theme_classic()

theme_void()

theme_minimal()

---

```{r}
plot0 <- ggplot(gapminder_2007) + 
  geom_point(aes(x=gdpPercap, y=lifeExp, color=continent)) +
  labs(x= "GDP per capita", y="Life Expectancy", color="Continent", 
       title = "Income vs Longevity", caption = "Source: Gapminder dataset")
```

---

```{r}
plot0
```

---

```{r}
plot0 + theme_bw()
```

---

```{r}
plot0 + theme_classic()
```

---

```{r}
plot0 + theme_void()
```

---

```{r}
plot0 + theme_minimal()
```

---

### Data visualization cheat sheet

RStudio - Help - Cheatsheets - Data visualizaion with ggplot2

Try the examples in Geoms - continuous function

---

```{r}
i <- ggplot(economics, aes(date, unemploy))
i + geom_area()
```

---

```{r}
i + geom_line()
```

