---
title: "ggplot2 - 3"
subtitle: "Geometries"
author: "Sanghoon Lee"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  xaringan::moon_reader:
    css: [default, c407_slides.css]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, comment=NA, fig.asp = 0.56, out.width='70%', dpi = 300, fig.align = 'center')
options(tibble.print_max = Inf)
library(kableExtra)
```

Scatter plot

- Overplotting - alpha, jitter. 
- labeling points

Bar plots

Histogram

line chart

---

```{r}
library(tidyverse)
``` 

---

### Scatter plot

In the last class, we learned how to deal with too many points using alpha aesthetic, limits arguments in scale functions, and log scales - [Link](https://sleeubc.github.io/c407_2020/Lecture_slides/11_ggplot2_2.html#28)

---
 
### jitter

```{r}
tb <- tibble(x= c(5,5,5,5,5, 3,3,3, 1), y= c(5,5,5,5,5, 3,3,3, 1))
```

```{r}
tb %>% glimpse
```

---

```{r}
ggplot(tb, aes(x,y)) + geom_point()
```

---

```{r}
ggplot(tb, aes(x,y)) + geom_point(position = "jitter")
```

---

```{r}
ggplot(tb, aes(x,y)) + geom_point(color="grey", size = 2) + 
  geom_point(position = "jitter")
```

---

```{r}
ggplot(tb, aes(x,y)) + geom_point(color="grey", size = 2) + 
  geom_point(position = position_jitter(width = 0.1))
```

---

```{r}
ggplot(tb, aes(x,y)) + geom_point(color="grey", size = 2) + 
  geom_point(position = position_jitter(width = 0.1, height = 0.1))
```

---

```{r}
ggplot(tb, aes(x,y)) + geom_point(color="grey", size = 2) + 
  geom_point(position = position_jitter(width = 0.2, height = 0.2))
```

---

```{r}
ggplot(tb, aes(x,y)) + geom_point(color="grey", size = 2) + 
  geom_jitter(width = 0.2, height = 0.2)
```

---

### Labels

```{r}
library(gapminder)
gapminder_2007 <- gapminder %>% filter(year == 2007)
```

---

```{r}
ggplot(gapminder_2007) + geom_point(aes(gdpPercap, lifeExp))
```

---

```{r}
ggplot(gapminder_2007) + geom_point(aes(gdpPercap, lifeExp)) + 
  scale_x_log10()
```

---

```{r}
ggplot(gapminder_2007, aes(gdpPercap, lifeExp)) + 
  geom_point() + geom_text(aes(label=country))+ scale_x_log10()
```

---

```{r}
gapminder_2007_big <- gapminder_2007 %>% top_n(20, pop)
```

```{r}
ggplot(gapminder_2007, aes(gdpPercap, lifeExp)) + geom_point() + 
  geom_text(aes(label=country), gapminder_2007_big)+ scale_x_log10()
```

---

```{r}
ggplot(gapminder_2007, aes(gdpPercap, lifeExp)) + geom_point() + 
  geom_label(aes(label=country), gapminder_2007_big)+ scale_x_log10()
```

---

```{r}
ggplot(gapminder_2007, aes(gdpPercap, lifeExp)) + geom_point() + 
  ggrepel::geom_text_repel(aes(label=country), gapminder_2007_big) + 
  scale_x_log10()
```

---

```{r}
ggplot(gapminder_2007, aes(gdpPercap, lifeExp)) + geom_point() + 
  ggrepel::geom_label_repel(aes(label=country), gapminder_2007_big) + 
  scale_x_log10()
```

---

### Barchart

```{r}
bca_2017 <- read_rds("Data/bca_sm.rds") %>% 
  filter(lubridate::year(sale_date) == 2017)
```

---

```{r}
aa_n <- bca_2017 %>% count(aa_label)
```

```{r}
aa_n
```

---

```{r}
ggplot(aa_n) + geom_col(aes(aa_label, n))
```

---

```{r}
ggplot(bca_2017) + geom_bar(aes(aa_label))
```

---

```{r}
ggplot(bca_2017) + geom_bar(aes(aa_label)) + coord_flip()
```

---

```{r}
ggplot(bca_2017) + geom_bar(aes(aa_label, fill=act_use_cat)) + 
  coord_flip()
```

---

```{r}
ggplot(bca_2017) + 
  geom_bar(aes(aa_label, fill=act_use_cat), position = "dodge") + 
  coord_flip()
```

---

```{r}
ggplot(bca_2017) + 
  geom_bar(aes(aa_label, fill=act_use_cat), position = "fill") + coord_flip()
```

---

### Histogram

```{r}
ggplot(bca_2017) + geom_histogram(aes(price))
```

---

```{r}
ggplot(bca_2017) + geom_histogram(aes(price)) + 
  scale_x_continuous(limits = c(0,6000000))
```

---

```{r}
ggplot(bca_2017) + 
  geom_histogram(aes(price), binwidth = 1000) + 
  scale_x_continuous(limits = c(0,6000000))
```

---

```{r}
ggplot(bca_2017) + 
  geom_histogram(aes(price), binwidth = 1000) + 
  scale_x_continuous(limits = c(400000,450000))
```

---

```{r}
ggplot(bca_2017) + 
  geom_histogram(aes(price), binwidth = 100000) + 
  scale_x_continuous(limits = c(0,6000000))
```

---

```{r}
ggplot(bca_2017) + 
  geom_histogram(aes(price, fill=act_use_cat), binwidth = 100000) + 
  scale_x_continuous(limits = c(0,6000000))
```

---

.small[
```{r}
ggplot(bca_2017) + 
  geom_histogram(aes(price, fill=act_use_cat), binwidth = 100000, position = "dodge") + 
  scale_x_continuous(limits = c(0,6000000))
```
]

---

.small[
```{r}
ggplot(bca_2017) + 
  geom_histogram(aes(price, fill=act_use_cat), binwidth = 100000, position = "fill") + 
  scale_x_continuous(limits = c(0,10000000))
```
]

---

.small[
```{r}
ggplot(bca_2017) + 
  geom_histogram(aes(price, fill=act_use_cat), binwidth = 1000000, position = "fill") + 
  scale_x_continuous(limits = c(0,10000000))
```
]

---

```{r}
ggplot(bca_2017) + 
  geom_histogram(aes(price, fill=aa_label), binwidth = 100000) + 
  scale_x_continuous(limits = c(0,6000000))
```

---

"dodge" does not work well with many groups.

.small[
```{r}
ggplot(bca_2017) + 
  geom_histogram(aes(price, fill=aa_label), binwidth = 100000, position = "dodge") + 
  scale_x_continuous(limits = c(0,4000000))
```
]

---

.small[
```{r}
ggplot(bca_2017) + 
  geom_histogram(aes(price, fill=aa_label), binwidth = 1000000, position = "fill") + 
  scale_x_continuous(limits = c(0,10000000))
```
]

---

### line plots

We covered line plots in the last class [link](https://sleeubc.github.io/c407_2020/Lecture_slides/11_ggplot2_2.html#18).