---
title: "ggplot2 - 4"
subtitle: "Theme"
author: "Sanghoon Lee"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  xaringan::moon_reader:
    css: [default, c407_slides.css]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, comment=NA, fig.asp = 0.56, out.width='70%', dpi = 300, fig.align = 'center')
options(tibble.print_max = Inf)
library(kableExtra)
```

- geom_segment

- lollipop chart

- Built-in themes

- factor

The following will be covered later:

- theme - legend.position

- facet_wrap

- facet_grid

---

```{r}
library(tidyverse)
```

---

### geom_segment

.pull-left-1[
.small[
```{r}
tb <- tibble(
	x0=c(0,1), 
	y0=c(0,1), 
	x_chr=c(2,2), 
	y1=c(1,0))
tb
```
]]

.pull-right-2[
.tiny[
```{r}
ggplot(tb) + 
	geom_segment(aes(x=x0,y=y0,xend=x_chr, yend=y1))
```
]]

---

.pull-left-1[
.small[
```{r}
tb
```
]]

.pull-right-2[
.small[
```{r}
ggplot(tb, aes(x0, y0)) + 
	geom_point(size = 3) +
	geom_segment(aes(xend=x_chr, yend=y1))
```
]]


---

```{r}
bca_2017 <- read_rds("data/bca_sm.rds") %>% 
	filter(lubridate::year(sale_date) == 2017)
```

---

Write codes to generate avg_p below, from bca_2017.

```{r echo=FALSE}
avg_p <- bca_2017 %>% group_by(juris_label) %>% summarise(avg_p = mean(price), n=n())
```

.small[
```{r}
avg_p
```
]

---

```{r eval=FALSE}
avg_p <- bca_2017 %>% 
	group_by(juris_label) %>% 
	summarise(avg_p = mean(price), n=n())
```

---

Write codes to generate the following figure, using avg_p.

```{r echo=FALSE}
ggplot(avg_p, aes(juris_label, avg_p)) + 
	geom_point(color="red", size=3, alpha=0.8)
```

---

```{r eval=FALSE}
ggplot(avg_p, aes(juris_label, avg_p)) + 
	geom_point(color="red", size=3, alpha=0.8)
```

---

Revise the previous codes to generate the following figure.

```{r echo=FALSE}
ggplot(avg_p, aes(juris_label, avg_p)) + 
	geom_point(color="red", size=3, alpha=0.8) + 
	coord_flip()
```

---

```{r eval=FALSE}
ggplot(avg_p, aes(juris_label, avg_p)) + 
	geom_point(color="red", size=3, alpha=0.8) + 
	coord_flip()
```

---

Revise the previous codes to generate the following figure, using geom_segment().

```{r echo=FALSE}
ggplot(avg_p, aes(juris_label, avg_p)) + 
	geom_segment(aes(xend = juris_label, yend=0)) +
	geom_point(color="red", size=3, alpha=0.8) + 
	coord_flip()
```

---

```{r eval=FALSE}
ggplot(avg_p, aes(juris_label, avg_p)) + 
	geom_segment(aes(xend = juris_label, yend=0)) +
	geom_point(color="red", size=3, alpha=0.8) + 
	coord_flip()
```

---

Revise the previous codes to generate the following figure.

```{r echo=FALSE}
ggplot(avg_p, aes(juris_label, avg_p)) + 
	geom_segment(aes(xend = juris_label, yend=0)) +
	geom_point(color="red", size=3, alpha=0.8) + 
	labs(x = "", y="Average Housing Price (2007)") +
	coord_flip()
```

---

```{r eval=FALSE}
ggplot(avg_p, aes(juris_label, avg_p)) + 
	geom_segment(aes(xend = juris_label, yend=0)) +
	geom_point(color="red", size=3, alpha=0.8) + 
	labs(x = "", y="Average Housing Price (2007)") +
	coord_flip()
```

---

Revise the previous codes to generate the following figure.

```{r echo=FALSE}
avg_p %>% filter(!is.na(juris_label)) %>% 
ggplot(aes(juris_label, avg_p)) + 
	geom_segment(aes(xend = juris_label, yend=0)) +
	geom_point(color="red", size=3, alpha=0.8) + 
	labs(x = "", y="Average Housing Price (2007)") +
	coord_flip()
```

---

```{r eval=FALSE}
avg_p %>% filter(!is.na(juris_label)) %>% 
ggplot(aes(juris_label, avg_p)) + 
	geom_segment(aes(xend = juris_label, yend=0)) +
	geom_point(color="red", size=3, alpha=0.8) + 
	labs(x = "", y="Average Housing Price (2007)") +
	coord_flip()
```

---

Revise the previous codes to generate the following figure.

```{r echo=FALSE}
avg_p %>% filter(!is.na(juris_label)) %>% 
ggplot(aes(juris_label, avg_p)) + 
	geom_segment(aes(xend = juris_label, yend=0)) +
	geom_point(color="red", size=3, alpha=0.8) + 
	labs(x = "", y="Average Housing Price (2007)") +
	coord_flip() +
	theme_bw()
```

---

```{r eval=FALSE}
avg_p %>% filter(!is.na(juris_label)) %>% 
ggplot(aes(juris_label, avg_p)) + 
	geom_segment(aes(xend = juris_label, yend=0)) +
	geom_point(color="red", size=3, alpha=0.8) + 
	labs(x = "", y="Average Housing Price (2007)") +
	coord_flip() +
	theme_bw()
```

---

We want to generate the following figure. How?

```{r echo=FALSE}
avg_p %>% filter(!is.na(juris_label)) %>% 
ggplot(aes(juris_label, avg_p)) + 
	geom_segment(aes(xend = juris_label, yend=0)) +
	geom_point(color="red", size=3, alpha=0.8) + 
	labs(x = "", y="Average Housing Price (2007)") +
	scale_y_continuous(labels=scales::label_comma(scale=0.000001, prefix="$", suffix='M')) +
	coord_flip() +
	theme_bw()
```

---

.tiny[
```{r eval=FALSE}
avg_p %>% filter(!is.na(juris_label)) %>% 
ggplot(aes(juris_label, avg_p)) + 
	geom_segment(aes(xend = juris_label, yend=0)) +
	geom_point(color="red", size=3, alpha=0.8) + 
	labs(x = "", y="Average Housing Price (2007)") +
	scale_y_continuous( labels=scales::label_comma(scale=0.000001, prefix="$", suffix='M') ) +
	coord_flip() +
	theme_bw()
```
]

How does this work?

---

```{r}
f <- scales::label_comma(scale=0.000001, prefix="$", suffix='M')
```

```{r}
f(20000000)
```

.tiny[
```{r eval=FALSE}
avg_p %>% filter(!is.na(juris_label)) %>% 
ggplot(aes(juris_label, avg_p)) + 
	geom_segment(aes(xend = juris_label, yend=0)) +
	geom_point(color="red", size=3, alpha=0.8) + 
	labs(x = "", y="Average Housing Price (2007)") +
	scale_y_continuous( labels= f ) +
	coord_flip() +
	theme_bw()
```
]

---

```{r echo=FALSE}
avg_p %>% filter(!is.na(juris_label)) %>% 
ggplot(aes(juris_label, avg_p)) + 
	geom_segment(aes(xend = juris_label, yend=0)) +
	geom_point(color="red", size=3, alpha=0.8) + 
	labs(x = "", y="Average Housing Price (2007)") +
	scale_y_continuous( labels= f ) +
	coord_flip() +
	theme_bw()
```

---

We want to generate the following figure. How?

```{r echo=FALSE}
avg_p %>% filter(!is.na(juris_label)) %>% 
ggplot(aes(fct_reorder(juris_label, avg_p), avg_p)) + 
	geom_segment(aes(xend = juris_label, yend=0)) +
	geom_point(color="red", size=3, alpha=0.8) + 
	labs(x = "", y="Average Housing Price (2007)") +
	scale_y_continuous( labels=scales::label_comma(scale=0.000001, prefix="$", suffix='M') ) +
	coord_flip() +
	theme_bw()
```

---

.tiny[
```{r eval=FALSE}
avg_p %>% filter(!is.na(juris_label)) %>% 
ggplot(aes(fct_reorder(juris_label, avg_p), avg_p)) + 
	geom_segment(aes(xend = juris_label, yend=0)) +
	geom_point(color="red", size=3, alpha=0.8) + 
	labs(x = "", y="Average Housing Price (2007)") +
	scale_y_continuous( labels=scales::label_comma(scale=0.000001, prefix="$", suffix='M') ) +
	coord_flip() +
	theme_bw()
```
]

---

### Factor

```{r}
x <- factor(c("a", "b", "c"), levels=c("c", "b", "a"))
```

```{r}
x
```
```{r}
x %>% as.numeric()
```

---

```{r}
tb <- tibble(x, x_chr=c("a", "b", "c"), y=c(3,1,2))
```

```{r}
tb
```

---

```{r}
ggplot(tb) + geom_point(aes(x,y))
```

---

```{r}
ggplot(tb) + geom_point(aes(x_chr,y))
```

---

```{r}
tb <- tb %>% mutate(x2=fct_reorder(x, y))
```

```{r}
tb
```

---

```{r}
tb$x
```
```{r}
tb$x2
```

---

```{r}
ggplot(tb) + geom_point(aes(x2,y))
```
