---
title: "Basic mapping with ggmap"
author: "Sanghoon Lee"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  xaringan::moon_reader:
    css: [default, c407_slides.css]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, comment=NA, fig.asp = 0.56, out.width='70%', dpi = 300, fig.align = 'center')
options(tibble.print_max = Inf)
library(kableExtra)
```

Google API registration

geocode

get_map

ggmap

baselayer

facet_wrap

---

### Google API registration

1. Go to https://cloud.google.com/maps-platform/
1. Press 'GET STARTED' button.
1. Sign-in or create account.
1. Create a project - COMM 407
1. Create billing account. You will need credit card information.
1. Pick products - check all.
1. Enable Google Maps Platform
1. Save your API key

---

```{r}
library(ggmap)
library(tidyverse)
```

```{r eval=FALSE}
register_google("Your API key here")
```

```{r echo=FALSE}
register_google("AIzaSyB7XZzRqFHPanx4ph26Rvmo7Qg_P1AHvHM")
```

---

### Geocoding

.small[
```{r}
tb <- tibble(
  address = c("2053 main mall Vancouver", "Pacific Poke UBC", "Chen centre"))

tb
```
]

---

```{r cache=TRUE}
tb <- tb %>% mutate_geocode(address)

tb
```

---

```{r}
ggplot(tb) + geom_point(aes(lon, lat))
```

```{r cache=TRUE}
ubcmap <- get_map("University of British Columbia Vancouver")
```

---

```{r}
ggmap(ubcmap)
```

---

```{r cache=TRUE}
ubcmap <- get_map("Sauder School of Business", zoom=15)
```

```{r}
ggmap(ubcmap) + geom_point(aes(lon, lat), tb, color="red", size=3)
```

---

```{r}
ggmap(ubcmap) + geom_point(aes(lon, lat), tb, color="red", size=3) + 
  ggrepel::geom_text_repel(aes(lon, lat, label=address), tb)
```


---

### base_layer

```{r}
ggmap(ubcmap, base_layer = ggplot(tb, aes(lon, lat))) + 
  geom_point(color="red", size=3) + 
  ggrepel::geom_text_repel(aes(label=address))
```

---

### Map type

```{r eval=FALSE}
? get_map
```

get_map(location = c(lon = -95.3632715, lat = 29.7632836),
  zoom = "auto", scale = "auto", maptype = c("terrain",
  "terrain-background", "satellite", "roadmap", "hybrid", "toner",
  "watercolor", "terrain-labels", "terrain-lines", "toner-2010",
  "toner-2011", "toner-background", "toner-hybrid", "toner-labels",
  "toner-lines", "toner-lite"), source = c("google", "osm", "stamen"),
  force = ifelse(source == "google", TRUE, FALSE), messaging = FALSE,
  urlonly = FALSE, filename = NULL, crop = TRUE, color = c("color",
  "bw"), language = "en-EN", ...)

---

.tiny[
```{r cache=TRUE}
ubcmap_2 <- get_map("Sauder School of Business", zoom=15, maptype="roadmap")
ggmap(ubcmap_2)
```
]

---

```{r}
ggmap(ubcmap_2, base_layer = ggplot(tb, aes(lon, lat))) + 
  geom_point(color="red", size=3) + 
  ggrepel::geom_text_repel(aes(label=address))
```

---

.small[
```{r cache=TRUE}
ubcmap_stamen <- get_map("Sauder School of Business", 
                         zoom=15, source="stamen")
ggmap(ubcmap_stamen)
```
]

---

```{r}
ggmap(ubcmap_stamen, base_layer = ggplot(tb, aes(lon, lat))) + 
  geom_point(color="red", size=3) + 
  ggrepel::geom_text_repel(aes(label=address))
```

---

.small[
```{r cache=TRUE}
ubcmap_stamen_2 <- get_map("Sauder School of Business", zoom=15, 
                           source="stamen", maptype = "toner-lite")
ggmap(ubcmap_stamen_2)
```
]

---

```{r}
ggmap(ubcmap_stamen_2, base_layer = ggplot(tb, aes(lon, lat))) + 
  geom_point(color="red", size=3) + 
  ggrepel::geom_text_repel(aes(label=address))
```

---

### theme_void()

.small[
```{r}
ggmap(ubcmap_stamen_2, base_layer = ggplot(tb, aes(lon, lat))) + 
  geom_point(color="red", size=3) + 
  ggrepel::geom_text_repel(aes(label=address)) + theme_void()
```
]

---

```{r cache=TRUE}
bca_2018_7 <- read_rds("Data/bca_sm.rds") %>% 
  filter(lubridate::year(sale_date) == 2018) %>% 
  filter(lubridate::month(sale_date) == 7)
```

---

```{r}
ggplot(bca_2018_7) + geom_point(aes(lon, lat))
```

---

```{r cache=TRUE}
vanmap <- get_map("Vancouver Canada")
ggmap(vanmap) 
```

---

```{r}
vanmap <- get_map("Surrey Canada")
ggmap(vanmap) + geom_point(aes(lon, lat), bca_2018_7) 
```

---

```{r}
ggmap(vanmap) + geom_point(aes(lon, lat, color=price), bca_2018_7) 
```

---

.small[
```{r}
ggmap(vanmap) + geom_point(aes(lon, lat, color=price), bca_2018_7) +
  scale_color_viridis_c()
```
]

---

.small[
```{r cache=TRUE}
vanmap <- get_map("Surrey Canada", source = "stamen", maptype = "toner-lite")

ggmap(vanmap) + geom_point(aes(lon, lat, color=price), bca_2018_7, alpha=0.5) + 
  scale_color_viridis_c()
```
]

---

.small[
```{r}
ggmap(vanmap) + geom_point(aes(lon, lat, color=price), bca_2018_7, alpha=0.5) + 
  scale_color_viridis_c() + theme_void()
```
]

---

.small[
```{r}
ggmap(vanmap) + geom_point(aes(lon, lat, color=price), bca_2018_7, alpha=0.5) + 
  facet_wrap(~act_use_cat)+ scale_color_viridis_c() + theme_void()
```
]

---

.tiny[
```{r}
ggmap(vanmap) + geom_point(aes(lon, lat, color=2018-year_built), bca_2018_7, alpha=0.5) + 
  facet_wrap(~act_use_cat)+ scale_color_viridis_c("Age") + theme_void()
```
]

