---
title: "Tidy Data"
author: "Sanghoon Lee"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  xaringan::moon_reader:
    css: [default, c407_slides.css]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, comment=NA, fig.asp = 0.56, out.width='70%', dpi = 300, fig.align = 'center')
options(tibble.print_max = Inf)
library(kableExtra)
```

gather()

separate()

spread()

---

```{r}
library(tidyverse)
```

```{r}
wide <- tibble(
  type = c("a", "b"), 
  yr_2017 = c(1, 1), 
  yr_2018=c(1.2, 1.5), 
  yr_2019=c(1.2, 1.3))
```

---

.pull-left[
```{r}
wide
```
]

.pull-right[

You want

```{r echo=FALSE, out.width='110%'}
wide %>% 
  gather(key = year, value =price, -type) %>%  
  separate(year, into=c(NA, "year"), convert = TRUE) %>% 
  ggplot(aes(year, price, color=type)) + geom_line() + geom_point()
```
]

---

.pull-left[
```{r}
wide
```
]

.pull-right[

You need long form

```{r echo=FALSE, out.width='110%'}
wide %>% 
  gather(key = year, value =price, -type) %>%  
  separate(year, into=c(NA, "year"), convert = TRUE)
```
]

---

.pull-left[
```{r}
wide
```
]

.pull-right[
```{r}
wide %>% 
  gather(key = year, value =price)
```
]

---

.pull-left[
.small[
```{r}
wide
```
]]

.pull-right[
.small[
```{r}
wide %>% 
  gather(key=year, value=price, 
         yr_2017:yr_2019)
```
]]

---

.pull-left[
.small[
```{r}
wide
```
]]

.pull-right[
.small[
```{r}
wide %>% 
  gather(key=year, value=price, 
         -type)
```
]]

---

```{r}
wide %>% gather(key=year, value=price, -type) %>% 
  separate(year, into=c("item", "year"), sep="_")
```

---

```{r}
long <- wide %>% gather(key=year, value=price, -type) %>% 
  separate(year, into=c(NA, "year"), convert = TRUE)
long
```

---

### UK housing data

.small[
1. Go to [UK Office for National Statistics](https://www.ons.gov.uk/) (Google it.)

1. Click People, population and community - Housing.

1. Scroll down the web-page to a sub-header called "Datasets related to housing"

1. Click on the second listed data-set called "Residential property sales for administrative geographies: HPSSA dataset 6"

1. Go to "Your download option - current" and click 'xls (2.0 MB)'

1. Move the downloaded xls file into 'C407/Data' folder.

1. Browse data using MS-Excel.
]

---

.small[
```{r}
wide_1a <- readxl::read_excel(
  "Data/hpssadataset6numofresidentialpropertysalesforadmingeos.xls", 
  sheet = "1a", skip = 6)
```
]

```{r eval=FALSE}
wide_1a %>% View
```

---

.scroll-output[
.small[
```{r echo=FALSE}
wide_1a %>% kable()
```
]]

---

```{r}
wide_1a <- wide_1a %>% 
  filter(!is.na(Code)) %>% 
  select(-starts_with("..."))
```

```{r eval=FALSE}
wide_1a %>% View
```

---

.scroll-output[
.small[
```{r echo=FALSE}
wide_1a %>% kable()
```
]]

---

```{r}
long_1a <- wide_1a %>% 
  gather(key=year, value=num_sales, -Code, -Name)
```

```{r eval=FALSE}
long_1a %>% View
```

---

.scroll-output[
.small[
```{r echo=FALSE}
long_1a %>% kable()
```
]]

---

```{r}
long_1a <- long_1a %>% 
  separate(year, into = c(NA, "year"), sep = " ending ")
```


```{r eval=FALSE}
long_1a %>% View
```

---

.scroll-output[
.small[
```{r echo=FALSE}
long_1a %>% kable()
```
]]

---

```{r}
long_1a <- long_1a %>% 
  mutate(year = lubridate::myd(year, truncated = 1))
```


```{r eval=FALSE}
long_1a %>% View
```

---

.scroll-output[
.small[
```{r echo=FALSE}
long_1a %>% kable()
```
]]

---

```{r}
long_1a %>% filter(Code %>% str_detect("E12")) %>% 
  ggplot(aes(year, num_sales, color=Name)) + geom_line() + geom_point()
```

---

### spread()

```{r}
long
```

You want to calculate the price of b relative to a.

---

.pull-left[
```{r}
long
```
]

.pull-right[
```{r}
wide2 <- long %>% 
  spread(key=type, value=price)
wide2
```
]

---

.pull-left[
```{r}
long
```
]

.pull-right[
```{r}
wide2 %>% 
  mutate(ratio_ba = b/a)
```
]
