---
title: "Census profile practice"
author: "Sanghoon Lee"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  xaringan::moon_reader:
    css: [default, c407_slides.css]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, comment=NA, fig.asp = 0.56, out.width='70%', dpi = 300, fig.align = 'center')
options(tibble.print_max = Inf)
library(kableExtra)
```

```{r echo=FALSE}
library(tidyverse)

profile <- read_csv(
  "Data/98-401-X2016041_eng_CSV/98-401-X2016041_English_CSV_data.csv"
  )

members <- profile %>% 
  select(`DIM: Profile of Census Metropolitan Areas/Census Agglomerations (2247)`, 
         `Member ID: Profile of Census Metropolitan Areas/Census Agglomerations (2247)`) %>% 
  unique()
```

Calculate the share of households spending 30% or more of their incomes on housing, by CMA, and study how it is related to city size or housing value.

--

1. Identify relevant columns from StatCan website.

1. Find corresponding member IDs.

1. Filter data using the member IDs.

1. Use spread() to convert to wide form.

1. Calculate the household shares spending 30% or more of their incomes on housing.

--

Note you already have Census profile data across CMA and CAs (profile) and its member IDs (members) in R.


---

### Identify relevant columns from StatCan website.

Go to [Statistics Canada website](https://www12.statcan.gc.ca).

Click Census - Data products - Census Profile

Click View 'Canada' in the Browse box.

Find relevant columns.

---

- Population, 2016
- Total - Owner and tenant households with household total income greater than zero, in non-farm, non-reserve private dwellings by shelter-cost-to-income ratio - 25% sample data
- Spending 30% or more of income on shelter costs
- Median value of dwellings ($)

--

Find their member IDs.

--

- 1
- 1667
- 1669
- 1676

---

.scroll-output[
```{r echo=FALSE}
members %>% kable()
```
]

---

.small[
```{r}
tb <- profile %>% 
	filter(`Member ID: Profile of Census Metropolitan Areas/Census Agglomerations (2247)` %in% c(1, 1667, 1669, 1676))
```

```{r}
tb <- tb %>% select(
  GEO_NAME, 
  item=`DIM: Profile of Census Metropolitan Areas/Census Agglomerations (2247)`, 
  n =`Dim: Sex (3): Member ID: [1]: Total - Sex`)
```

```{r eval=FALSE}
tb %>% View
```
]

---

.scroll-output[
.tiny[
```{r echo=FALSE}
tb %>% kable()
```
]]

---

```{r}
tb2 <- tb %>% spread(key=item, value=n)
```

```{r eval=FALSE}
tb2 %>% View
```

---

.scroll-output[
.tiny[
```{r echo=FALSE}
tb2 %>% kable()
```
]]

---

.tiny[
```{r}
tb3 <- tb2 %>% transmute(
	GEO_NAME, 
	`Median value of dwellings ($)`, 
	`Population, 2016`, 
	Housing30_share = `Spending 30% or more of income on shelter costs`/ `Total -  Owner and tenant households with household total income greater than zero, in non-farm, non-reserve private dwellings by shelter-cost-to-income ratio - 25% sample data`)
```
]

```{r eval=FALSE}
tb3 %>% View
```

---

.scroll-output[
.tiny[
```{r echo=FALSE}
tb3 %>% kable()
```
]]

---

```{r echo=FALSE}
tb3_lbl <- tb3 %>% filter(`Population, 2016`>1000000)
```

.small[
```{r}
tb3 %>% 
ggplot(aes(`Population, 2016`, Housing30_share)) + 
  geom_point() + geom_smooth() + 
  ggrepel::geom_text_repel(aes(label=GEO_NAME), tb3_lbl)+
  scale_x_log10()
```
]

---

```{r}
tb3 %>% 
  ggplot(aes(`Median value of dwellings ($)`, Housing30_share)) + 
  geom_point(aes(size= `Population, 2016`)) + geom_smooth() + 
  ggrepel::geom_text_repel(aes(label=GEO_NAME), tb3_lbl)+
  scale_x_log10() + scale_size_continuous(guide=FALSE)
```
