---
title: "Spatial Data Analysis - 1"
author: "Sanghoon Lee"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  xaringan::moon_reader:
    css: [default, c407_slides.css]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, comment=NA, fig.asp = 0.56, out.width='70%', dpi = 300, fig.align = 'center')
options(tibble.print_max = Inf)
library(kableExtra)
```

Vector spatial data - simple feature - sf package

geom_sf()

CRS

coord_sf()

---

Download the following files into the Data folder.

- census_tract_van.rds - Census 2016 profile for census tracts in Vancouver CMA

- ct_bndry_van.rds - Census 2016 census tract boundaries for Vancouver CMA

- csd_bndry_van.rds - Census 2016 Census Sub-Division boundaries for Vancouver CMA

---

```{r}
library(sf)
library(tidyverse)
```

```{r}
census_tract_van <- read_rds("Data/census_tract_van.rds")
```

```{r eval=FALSE}
census_tract_van
```

---
.scroll-output[
```{r echo=FALSE}
census_tract_van %>% head(200) %>% kable()
```
]

---

Suppose we want to know 

- median value of dwelling for owner households
- median total household income 

We need to know their member IDs. 

Find corresponding data item names from Census 2016 profile (Google it.) - Canada webpage. 

---

.tiny[
```{r}
census_tract_van %>% 
  filter(`DIM: Profile of Census Tracts (2247)` %>% str_detect("Median value")) %>% 
  head()
```
]

---

.tiny[
```{r}
census_tract_van %>% filter(
  `DIM: Profile of Census Tracts (2247)` %>% str_detect("Median total income of households")
  ) %>% 
  head()
```
]

---

.small[
```{r}
tb <- census_tract_van %>% 
  filter(`Member ID: Profile of Census Tracts (2247)` %in% c(1676, 742)) %>% 
  select(-`Member ID: Profile of Census Tracts (2247)`, 
         -`Dim: Sex (3): Member ID: [2]: Male`, 
         -`Dim: Sex (3): Member ID: [3]: Female`)
```
]

```{r eval=FALSE}
tb %>% View
```

---

```{r echo=FALSE}
tb %>% kable()
```

---

```{r}
tb <- tb %>% spread(
  key=`DIM: Profile of Census Tracts (2247)`, 
  value = `Dim: Sex (3): Member ID: [1]: Total - Sex`)
```

```{r}
tb %>% head()
```

---

.tiny[
```{r}
tb <- tb %>% mutate(
  `Median total income of households in 2015 ($)` = as.numeric(`Median total income of households in 2015 ($)`), 
  `Median value of dwellings ($)` = as.numeric(`Median value of dwellings ($)`))
```
]

```{r}
tb %>% head()
```

---


Difficult to identify geographic location corresponding to `GEO_CODE (POR)` (e.g., 9330001.01). 

We need a census tract boundary map.

```{r}
ct_bndry_van <- read_rds("Data/ct_bndry_van.rds")
```

--

ct_bndry_van is a sf (simple features) object.

--

A sf object is similar to a tibble, but it has

1. geometry column.
1. CRS (Coordinate Reference System).

---



.small[
```{r}
ct_bndry_van %>% head()
```
]

---

```{r}
ggplot(ct_bndry_van) + geom_sf()
```

---

.small[
```{r}
tb_sf <- ct_bndry_van %>% left_join(tb, by=c("CTUID"= "GEO_CODE (POR)"))
```
]

.scroll-output[
.tiny[
```{r}
tb_sf %>% head()
```
]]

---

.small[
```{r}
ggplot() + 
  geom_sf(aes(fill = `Median total income of households in 2015 ($)`), tb_sf)
```
]

---

.small[
```{r}
ggplot() + 
  geom_sf(aes(fill = `Median total income of households in 2015 ($)`), tb_sf) +
  scale_fill_viridis_c("Median\nHousehold\nIncome($)") 
```
]

---

```{r}
ggplot() + 
  geom_sf(aes(fill = `Median value of dwellings ($)`), tb_sf) +
  scale_fill_viridis_c("Median\nHousing\nValue($)") 
```

---

Calculate the housing price to income ratio by census tract, and show it on the map.

--

You can use dplyr functions on a sf object.

.tiny[
```{r}
tb_sf <- tb_sf %>% mutate(
  `Price to income ratio` = 
    `Median value of dwellings ($)`/`Median total income of households in 2015 ($)`
  )
```
]

---

```{r}
ggplot() + 
  geom_sf(aes(fill = `Price to income ratio`), tb_sf) +
  scale_fill_viridis_c("Price to\nIncome Ratio") 
```

---

### Overlaying municipality boundaries

```{r}
csd_bndry_van <- read_rds("Data/csd_bndry_van.rds")
```

.scroll-output[
.tiny[
```{r}
csd_bndry_van %>% head()
```
]]

---

```{r}
ggplot() + 
  geom_sf(aes(fill = `Price to income ratio`), tb_sf) +
  geom_sf(data=csd_bndry_van, color="white") +
  scale_fill_viridis_c("Price to\nIncome Ratio") 
```

---

```{r}
ggplot() + 
  geom_sf(aes(fill = `Price to income ratio`), tb_sf) +
  geom_sf(data=csd_bndry_van, color="white", fill=NA) +
  scale_fill_viridis_c("Price to\nIncome Ratio") 
```


---

### CRS (Coordinate Reference System)

.tiny[
```{r}
ggplot() + geom_sf(aes(fill = `Price to income ratio`), tb_sf) +
  geom_sf(data=csd_bndry_van, color="white", fill=NA) +
  scale_fill_viridis_c("Price to\nIncome Ratio") +
  coord_sf(crs=4326)
```
]

---

The 4326 is an EPSG (European Petroleum Survey Group) code for WGS84 geographic coordinate system. We use it for small geographic area (e.g., within CMA). Google Map uses the same coordinate system.

For a larger area (e.g., whole Canada) it would create a large distortion, and we would use Lambert conformal conic projection (EPSG code = 3347) See [StatCan Map Projection](https://www150.statcan.gc.ca/n1/pub/92-195-x/2011001/other-autre/mapproj-projcarte/m-c-eng.htm).

coord_sf() applys a projection to a map.

---

.tiny[
```{r}
ggplot(tb_sf) + geom_sf(aes(fill = `Price to income ratio`)) +
  geom_sf(data=csd_bndry_van, color="white", fill=NA) +
  scale_fill_viridis_c("Price to\nIncome Ratio") +
  coord_sf(crs=4326) + 
  theme_void()
```
]

---

### Saving the map

ggsave() saves the last ggplot figure as a file.

```{r}
ggsave("Temp/p_inc.png")
```

