---
title: "Spatial Join - Zoning"
author: "Sanghoon Lee"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  xaringan::moon_reader:
    css: [default, c407_slides.css]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, comment=NA, fig.asp = 0.56, out.width='70%', dpi = 300, fig.align = 'center')
options(tibble.print_max = Inf)
library(kableExtra)
```

Where to obtain spatial data - shapefile, lon and lat

st_read()

st_as_sf()

st_crs()

st_transform()

st_join()

---

```{r}
library(sf)
library(tidyverse)
```

---

### Shapefile

Most spatial data are available in shapefile format.

Go to [Open Data Portal Vancouver](https://opendata.vancouver.ca/pages/home/)

Search 'Zoning' - Choose 'Zoning districts and labels'

Click Export - Shapefile - Whole Dataset (Note WGS84 (EPSG 4326)).

Extract the zip file into the Data folder.

Note that a shapefile is a group of files sharing the same name but having different file extensions. You need all of them.

---

Use st_read() to read a shapefile and [its auxiliary files](https://gisgeography.com/arcgis-shapefile-files-types-extensions/) into R.

.tiny[
```{r}
zoning <- st_read("Data/zoning-districts-and-labels/zoning-districts-and-labels.shp")
```
]

---

.tiny[
```{r}
zoning %>% head()
```
]

---
```{r}
ggplot() + geom_sf(aes(fill=category), zoning)
```

---

### Lon lat

You can convert a tibble with longitude and latitude into a sf object.

.small[
```{r}
bca <- read_rds("Data/bca_sm.rds") %>% 
  filter(juris_id==200) %>% # city of Vancouver
  sample_frac(0.01) %>% # We use 1% random sample.
  select(aa_id, roll_num, lon, lat)
```
]

---

.small[
```{r}
bca %>% head()
```
]

---

You can use st_as_sf() to convert a tibble with lon and lat to a sf object.

When you convert lon-lat data, use crs=4326 which uses lon and lat as x and y coordinates.

```{r}
bca_sf <- bca %>% st_as_sf(coords=c("lon", "lat"), crs=4326)
```

---

.scroll-output[
.small[
```{r}
bca_sf %>% head()
```
]]

---

.small[
```{r cache=TRUE}
ggplot() + geom_sf(data=bca_sf)
```
]

---

### Spatial Join

Suppose that you want to identify zoning code for each transaction in the city of Vancouver.

You can spatial join the two sf objects: bca_sf and zoning

---

For spatial join (or any other geoprocessing) to work, the two sf objects must have the same CRS. 

.small[
```{r}
st_crs(bca_sf)
```
]

.small[
```{r}
st_crs(zoning)
```
]

---

Otherwise we would have to change the CRS of one sf to match that of the other sf, using st_transform() function.

```{r eval=FALSE}
zoning <- zoning %>% st_transform(crs = st_crs(bca_sf) )
```

It is not necessary in this case, as the two already use the same CRS.

---

```{r}
bca_z <- bca_sf %>% st_join( zoning )
```

st_join() keeps all rows in the first sf dataset, just like the left_join().

---

.small[
```{r}
bca_z
```
]

---

### Dropping geometry column

```{r}
bca_z <- bca_z %>% st_drop_geometry()
bca_z %>% head()
```

