---
title: "Buffer Analysis - Practice"
author: "Sanghoon Lee"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  xaringan::moon_reader:
    css: [default, c407_slides.css]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, comment=NA, fig.asp = 0.56, out.width='70%', dpi = 300, fig.align = 'center')
options(tibble.print_max = Inf)
library(kableExtra)
set.seed(0)
```

Find all housing transactions within 1km from Canada Line.

---

```{r}
library(sf)
library(tidyverse)
```

```{r}
CanLine_Stn <- st_read("Data/CanadaLine/CanLine_Stn.shp")
```

---

```{r}
CanLine_Stn
```

---

```{r}
CanLine_Stn <- CanLine_Stn %>% select(Station)
```

```{r}
CanLine_Stn
```

---

```{r}
ggplot() + geom_sf(data=CanLine_Stn) + geom_sf_text(aes(label=Station), CanLine_Stn, size=3)
```

---

```{r}
bca <- read_rds("Data/bca_sm.rds") %>% 
	sample_frac(0.01) %>% 
	select(aa_id, roll_num, lon, lat)
```

```{r}
bca_sf <- bca %>% st_as_sf(coords= c("lon", "lat"), crs=4326)
```

---

```{r}
ggplot() + 
	geom_sf(data=bca_sf, color="grey") + 
	geom_sf(data=CanLine_Stn, color="red")
```

---

We want to find all housing transactions that are within 1km from a Canada line station.

We will create buffers of 1000 meters from Canada line stations.

We have to make sure that the CRS for CanLine_Stn uses meter as distance unit.

```{r}
st_crs(CanLine_Stn)
```

If not, you have to convert CRS using st_transform(crs=26910)

---

```{r}
CanLine_Stn_buffer <- CanLine_Stn %>% st_buffer(dist = 1000)
```

---

```{r}
ggplot() +
	geom_sf(data=bca_sf, color="grey") +
	geom_sf(data=CanLine_Stn, color="red") +
	geom_sf(data=CanLine_Stn_buffer, color="blue", fill=NA) +
	coord_sf(crs=4326, xlim = c(-123.2, -123), ylim=c(49.15, 49.3))
```

---

### Spatial Join

```{r}
CanLine_Stn_buffer <- CanLine_Stn_buffer %>% 
	st_transform(crs=st_crs(bca_sf))
```

```{r}
bca_stn <- bca_sf %>% st_join(CanLine_Stn_buffer)
```

---

.tiny[
```{r}
bca_stn
```
]

---

```{r}
ggplot() +
	geom_sf(aes(color=Station), data=bca_stn, alpha=0.3) +
	scale_color_discrete(guide=FALSE) +
	coord_sf(crs=4326, xlim = c(-123.2, -123), ylim=c(49.15, 49.3))
```

---

```{r}
bca_stn <- bca_stn %>% st_drop_geometry()
```

---

```{r}
bca_stn %>% count(Station, sort=TRUE)
```

---

Transactions that are within 1km of any station on Millenium line and Expo line.

```{r}
bca_stn %>% select(-Station) %>% unique()
```

