---
title: "Buffer Analysis"
author: "Sanghoon Lee"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  xaringan::moon_reader:
    css: [default, c407_slides.css]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, comment=NA, fig.asp = 0.56, out.width='70%', dpi = 300, fig.align = 'center')
options(tibble.print_max = Inf)
library(kableExtra)
set.seed(0)
```

Suppose you want to find all housing transactions within 1km from each skytrain station on Millenium line and Expo line.

1. Obtain skytrain staion locations. Load them into R. 
1. Load housing transaction locations into R.
1. Create 1km buffers from the skytrain station locations.
1. Spatial join the two datasets.

---

Go to [Abacus Dataverse Network website](http://dvn.library.ubc.ca/). (Google "Abacus UBC")

Click "Log In"

Choose "UBC" in affiliated user login, and click Login.

Search for "Skytrain"

Click "TRANSLINK TRANSIT GIS DATA, 09 MARCH 2010"

Click "Data & Analysis" tab.

Download and unzip "SkyTrain" and "CanadaLine" data into the data folder. (We will use CanadaLine in a separate practice.)

---

```{r}
library(sf)
library(tidyverse)
```

```{r}
SkyTstation <- st_read("Data/SkyTrain/SkyTStation.shp")
```

---

```{r}
SkyTstation
```

---

```{r}
SkyTstation <- SkyTstation %>% select(PL_NAME)
```

```{r}
SkyTstation
```

---

```{r}
ggplot() + 
	geom_sf(data=SkyTstation) + 
	geom_sf_text(aes(label=PL_NAME), SkyTstation, size=3)
```

---

```{r}
bca <- read_rds("Data/bca_sm.rds") %>% 
	sample_frac(0.01) %>% 
	select(aa_id, roll_num, lon, lat)
```

```{r}
bca_sf <- bca %>% 
	st_as_sf(coords= c("lon", "lat"), crs=4326)
```

---

```{r}
ggplot() + 
	geom_sf(data=bca_sf, color="grey") + 
	geom_sf(data=SkyTstation, color="red")
```

---

We want to find all housing transactions that are within 1km from a skytrain station.

We will create buffers of 1000 meters from Skytrain stations.

We have to make sure that the CRS for SkyTstation uses meter as distance unit.

```{r}
st_crs(SkyTstation)
```

If not, you have to convert CRS using st_transform(crs=26910)

---

```{r}
SkyTstation_buffer <- SkyTstation %>% st_buffer(dist = 1000)
```

---

```{r}
ggplot() +
	geom_sf(data=bca_sf, color="grey") +
	geom_sf(data=SkyTstation, color="red") +
	geom_sf(data=SkyTstation_buffer, color="blue", fill=NA) +
	coord_sf(crs=4326, xlim = c(-123.15, -122.8), ylim=c(49.16, 49.3))
```

---

### Spatial Join

```{r}
SkyTstation_buffer <- SkyTstation_buffer %>% 
	st_transform(crs=st_crs(bca_sf))
```

```{r}
bca_stn <- bca_sf %>% st_join(SkyTstation_buffer)
```

---

.tiny[
```{r}
bca_stn
```
]

Some transactions appear multiple times when they are within 1km from multiple stations. 

---

```{r}
ggplot() +
	geom_sf(aes(color=PL_NAME), data=bca_stn, alpha=0.3) +
	scale_color_discrete(guide=FALSE) +
	coord_sf(xlim = c(-123.15, -122.8), ylim=c(49.16, 49.3))
```

---

```{r}
bca_stn <- bca_stn %>% st_drop_geometry()
```

---

```{r}
bca_stn %>% count(PL_NAME, sort=TRUE)
```

---

Transactions that are within 1km of any station on Millenium line and Expo line.

```{r}
bca_stn %>% filter(!is.na(PL_NAME)) %>% select(-PL_NAME) %>% unique()
```

