---
title: "Transaction volume vs Month"
author: "Sanghoon Lee"
date: "3/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

Introduction

We study the relationship between sale volume and month.

Data

We use BC Assessment housing transaction data. 

```{r}
bca <- read_rds("data/bca_sm.rds")
```

Analysis

We drop observations from 2018, since we have data only until September 2018.
```{r}
library(lubridate)
bca <- bca %>% filter(year(sale_date) != 2018)
```

Our quick data analysis shows that sale volume is highest in May.
```{r}
bca <- bca %>% 
	mutate(Month = month(sale_date, label = TRUE))

bca %>% count(Month)
```

```{r}
bca %>% ggplot() + geom_bar(aes(Month))
```

```{r}
bca %>% ggplot() + geom_bar(aes(Month, fill=act_use_cat), position = "fill")
```

Conclusion

The analysis shows that housing transaction volumes are higher from March to August. This is true for both single family detached houses and condos. The percentage of condos among housing transactions are higher from October to December.

