---
title: "Tidy Data"
author: "Sanghoon Lee"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  xaringan::moon_reader:
    css: [default, c407_slides.css]
    lib_dir: libs
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, comment=NA, fig.asp = 0.56, out.width='70%', dpi = 300, fig.align = 'center')
options(tibble.print_max = Inf)
library(kableExtra)
```

```{r}
library(tidyverse)
```

---

Housing expenditure share of total incomes across CMAs?

```{r}
pumf <- read_rds("Data/2016_census_pumf_ind.rds")
```

Find relevant variables Chapter 1 of the user guide.

--

SHELCO, HHINC

--

Look up these variables in Chapter 2.

---

### SHELCO

.inverse[Shelter cost refers to the average monthly total of all shelter expenses paid by households that own or rent their
dwelling.]
Shelter costs for owner households include, where applicable, mortgage payments, property taxes and
condominium fees, along with the costs of electricity, heat, water and other municipal services. For renter
households, shelter costs include, where applicable, the rent and the costs of electricity, heat, water and other
municipal services.
This variable is reported for private households in owner or tenant occupied dwellings that are not located on an
agricultural operation operated by a member of the household.
Value ranges from 0 to the maximum amount on the file.

---

.scroll-output[

### TOTINC

.small[
.inverse[Total income refers to receipts from certain sources (in cash and, in some circumstances, in kind), before income
taxes and deductions, during the reference period. The monetary receipts included are those that tend to be of a
regular and recurring nature.]

Receipts that are included as income are:
- employment income from wages, salaries, tips, commissions and net income from self-employment (for both
unincorporated farm and non-farm activities)
- income from investment sources, such as dividends and interest on bonds, accounts, guaranteed investment
certificates (GICs) and mutual funds
- income from employer and personal pension sources, such as private pensions and payments from annuities
and registered retirement income funds (RRIFs)
- other regular cash income, such as child support payments received, spousal support payments (alimony)
received and scholarships
- income from government sources, such as social assistance, child benefits, Employment Insurance benefits, Old
Age Security benefits, Canada Pension Plan and Quebec Pension Plan benefits and disability income.

Receipts excluded from this income definition are:
- one-time receipts, such as lottery winnings, gambling winnings, cash inheritances, lump sum insurance
settlements and tax-free savings account (TFSA) or registered retirement savings plan (RRSP) withdrawals
- capital gains because they are not by their nature regular and recurring. It is further assumed that they are more
relevant to the concept of wealth than the concept of income
- employer's contributions to registered pension plans, Canada Pension Plan, Quebec Pension Plan and
Employment Insurance
- voluntary inter-household transfers, imputed rent, goods and services produced for barter, and goods produced
for own consumption.

For 2016 Census, the reference period for this variable is calendar year 2015.

.inverse[
The value 88,888,888 stands for not available. The value 99,999,999 stands for not applicable and is applied to all
persons aged less than 15 years. Otherwise, this variable could be positive, negative or zero and is a rounded
value of the amount received by the individual in 2015. Values that would have been rounded to zero have been
replaced by 1 or -1. In some cases, high values have been top coded and low values have been bottom coded in
this file.
]]]

---

.scroll-output[

### Top coding

See section H.2 of the user guide.

.small[

H.2 Rounding and adjustment of extreme values for income, shelter costs, child care expenses, earnings and losses

In planning this microdata file, it was deemed essential that the procedures used make it impossible to
identify an individual in the population from their income or from any other quantitative data. To do this,
income, shelter costs, child care expenses, earnings and losses for the individuals selected to be part of
this microdata file were subjected to the following rounding and extreme values adjustment procedures.
These procedures minimize the impact on quality.

First, since a very large portion of the incomes in the file are from tax files, all values for these variables
require rounding. Some income variables were randomly rounded with a base of 100, namely INVST,
RETIR, CHDBN, CQPPB, GOVTI, GTRFS, OASGI and EICBN. Others were rounded with a base of 1,000:
TOTINC, WAGES, SEMPI, OTINC, TOTINC_AT, EMPIN, INCTAX and MRKINC. For the CAPGN variable,
a base of 100 was used for values between -5,000 and 5,000 and a base of 1,000 was used for other
values. If the value of any variable was higher than 100,000, the rounding base used was 10,000. The
rounding base for the VALUE variable was set to 10,000, while that for the CHLDC and SHELCO variables
was set to 100. If a value higher than 0 would have been rounded to 0 through the random procedure, the
value 1 was assigned instead. Similarly, when a value lower than 0 would have been rounded to 0, the
value -1 was assigned. This was done to maintain the applicability condition for income sources and to
preserve the same number of negative, null and positive values before and after random rounding. Since
the random rounding was done independently by variable, some relations between income sources are no
longer valid. However, this rounding technique maintains the statistical nature of the data.

Second, large income, earnings, expenses and shelter costs were top-coded to eliminate all possibility of
disclosure. Values for shelter costs (VALUE and SHELCO) were top-coded when they were higher than a
threshold. .inverse[This threshold is the weighted 90th percentile for an individual’s geographical region rounded 
down using the rounding bases given in the previous paragraph. The replacement value for the top-coded
values was set to the weighted average of the top-coded values for the geographical region.] The thresholds and the replacement values were calculated from the set of all Census long-form respondents. Thus,
calculating the weighted sum of all values of a particular variable in a given geographical region yields a
sum comparable to the sum one would obtain if no top-coding had been done on the data. For income,
earnings and expenses, values higher than the rounded down 99th weighted percentile for an individual’s
respective area and gender were top-coded. Some supplementary top-coding was necessary to eliminate
the possibility of residual disclosure. Also, some negative values, i.e., those lower than a threshold, were
down-coded using the standard method. The replacement value is the threshold.

]]


---

```{r}
pumf %>% filter(TOTINC < 80000000) %>% 
	group_by(CMA) %>% summarise(avg_shelco=weighted.mean(SHELCO, WEIGHT) * 12, avg_totinc=weighted.mean(TOTINC, WEIGHT)) %>% 
	mutate(shelco_share = avg_shelco/avg_totinc)
```

